---
title: "Group4_Project"
output:
  html_document: default
  word_document: default
date: "2022-11-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Packages

```{r message=FALSE}
# add packages you need for this assignment
library(ggplot2)
library(tidyverse)
library(tigerstats)
library(tibble)
library(dplyr)
library(stringr)
library(SmartEDA)
library(kernlab)
library(arules)
library(readxl)
library(readr)
library(ggpubr)
library(naniar)
library(MASS)
```

Import the `Student's Performance` data set and name it `Performance`.

```{r}
Performance <-read.csv("~/Downloads/StudentsPerformance.csv")
head(Performance)
```

* Check the number of observations and features using `dim`

```{r}
dim(Performance)
summary(Performance)
```

> Your answer

* Check for missing values

```{r}
# finds the location of missing values
which(is.na(Performance))

# finds the count of missing values 
sum(is.na(Performance))
```

> Your answer

# Creating a data frame
```{r}
mydata <- as.data.frame(Performance)
summary(Performance)
```

> Your answer

# Reading scores by Gender plot - Histogram

```{r}
hist(
  Performance$reading.score, col = "orange",
  #breaks = c(10,15,20,25,30,35,40,45,50,55), 
  main = "Histogram of Reading Score", 
  xlab= "Reading Score",
  )
```

> Your answer

# Writing scores by Gender plot - Boxplot

```{r}
b <- ggplot(mydata, aes(gender, writing.score, color = test.preparation.course))
b <- b + geom_boxplot()
b <- b + ggtitle("Writing scores by Gender Boxplot")
b <- b + xlab("Gender") + ylab("Writing Scores")
b
```
> Your answer

# Math scores by Gender plot - overlay histogram

```{r}
#have black lines around each bar
ggplot(data = Performance, aes(x = math.score)) +
geom_histogram(aes(fill = gender),bins = 10, color = "black")
#normalize the histogram
ggplot(data = Performance, aes(x = math.score)) +
geom_histogram(aes(fill = gender),bins = 10, color = "black", position = "fill")  
```

> Your answer

# Hypothesis testing

```{r}
random_sample <- Performance %>
sample_n(30)
head(random_sample, 10)
```

> Your answer

# Population and Sample Mean Comparison

```{r}
population_mean <- mean(Performance$math.score)
population_mean

sample_mean <- mean(random_sample$math.score)
sample_mean

mean_comparison <- data.frame("Type of Data" = c("Population", "Sample"), "Mean" = c(population_mean, sample_mean))
mean_comparison
```

> Your answer

# Comparison between population and Sample mean

```{r}
ggplot(mean_comparison, aes(x = Mean, y = Type.of.Data, fill = Type.of.Data)) + geom_col(width = 0.6) + ggtitle("Population vs. Sample Mean")
```

> Your answer

# T test - Hypothesis Testing

```{r}
t.test(random_sample$math.score, mu = mean(Performance$math.score))
```

> Your answer

# Two Sample T-Test (Divide Population into Two Parts)

```{r}
male_students <- Performance %>%
filter(gender == "male")

female_students <- Performance %>%
filter(gender == "female")

head(male_students, 5)

head(female_students, 5)
```

# Comparing The Measure of Variability Between Both Samples

```{r}

#Variance
var(male_students$reading.score)

var(female_students$reading.score)

#SD
sd(male_students$reading.score)

sd(female_students$reading.score)
```

# Visualizing The Distribution of Both Data

```{r}
ggplot(male_students, aes(x = reading.score, y = stat(density))) + geom_histogram() + geom_density() + ggtitle("Distribution of Male Students Reading Score")
```

> Your answer

```{r}
ggplot(female_students, aes(x = reading.score, y = stat(density))) + geom_histogram() + geom_density() + ggtitle("Distribution of Female Students Reading Score")
```

> Your answer

# Male and Female Students Reading Score Comparison

```{r}
reading_score_df <- data.frame("Male Reading Score" = mean(male_students$reading.score), "Female Reading Score" = mean(female_students$reading.score))
reading_score_df
```

> Your answer

# Comparison of Reading Score by Gender

```{r}
boxplot(Performance$reading.score ~ Performance$gender, col = c("red", "green"), main = "Comparison of Reading Score by Gender", xlab = "Gender", ylab = "Reading Score")
```

> Your answer

# Two sample t test

```{r}
t.test(reading.score ~ gender, data = Performance)
```

> Your answer


# Regression testing 

#Forward selection
```{r}
# Create a null model 
intercept_only <- lm(math.score ~ 1, data=Performance[,1:8])
# Create a full model
all <- lm(math.score~., data=Performance[,1:8])
# perform forward step-wise regression
forward <- stepAIC (intercept_only, direction='forward',scope = formula(all))
```

```{r}
ypred_forward <-predict(object = forward, newdata = Performance$math.score[1:10])
MAE(y_pred = ypred_forward, y_true = Performance$math.score)
MSE(y_pred = ypred_forward, y_true = Performance$math.score)
```

#backward selection

```{r}
backward <- stepAIC (all, direction='backward')
```

#Both direction

```{r}
both <- stepAIC (intercept_only, direction='both',scope = formula(all),trace = 0)
both$anova
```

```{r}
# Prediction models
randIndex <- sample(1:dim(mydata)[1])
#  # In order to split data, create a 2/3 cutpoint and round the number
cutpoint2_3 <- floor(2*dim(mydata)[1]/5)

#Use 80% of dataset as training set and remaining 20% as testing set
sample <- sample(c(TRUE, FALSE), nrow(Performance), replace=TRUE, prob=c(0.8,0.2))
train  <- Performance[sample, ]
test   <- Performance[!sample, ]

#------------------------------------------------------lm model
model <- lm(math.score ~ writing.score + gender + race.ethnicity + lunch + parental.level.of.education + test.preparation.course, data=test)
summary(model)
lmPred <- predict(model,test,interval = "prediction", level=0.95)
summary(lmPred)
head(lmPred)

# 1. Add predictions 
mydata1 <- cbind(test, lmPred)
head(mydata1)
# 2. Regression line + confidence intervals
p <- ggplot(mydata1, aes( fit, math.score)) +
  geom_point() +
  stat_smooth(method = lm)
# 3. Add prediction intervals
p + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
  geom_line(aes(y = upr), color = "red", linetype = "dashed") +
xlab("Predicted Scores") + ylab("Test Scores")
```